---
title: "Functions"
format:
  live-html: 
    toc: false
    section-divs: true
    include-after-body: 
      - text: |
          <script src="/js/progressive-reveal.js"></script>
          <script src="/js/bootstrapify.js"></script>

engine: knitr
#filters:
#  - webr
webr:
  cell-options:
    editor-font-scale: 0.85
    fig-width: 6
    fig-height: 3.7
    out-width: "70%"
---

{{< include /_extensions/r-wasm/live/_knitr.qmd >}}

```{r include=FALSE}
source(here::here("R", "helpers.R"))
```

### Functions {.no-hide}

First, read the [Posit Recipe on running functions](https://posit.cloud/learn/recipes/basics/BasicB4). 

```{r}
#| include: false
# Most functions produce a result like we see here, but some functions are designed to trigger side effects in the background. As a new R user you should know how to do three things with a function. First, you should know that if you type the name of a function by itself, R will return the code that is stored in the function.
# 
# This is the code that R runs when you call the function. The code may not mean anything to you now, but it can be useful down the road.
# 
# Second, you should know that to run a function you must do more than type its name. You must place a pair of parentheses behind the function name.
# 
# If the function requires any input to do its job you place the input between the parentheses, like the number three here.
# 
# Finally, you should know that if you put a question mark in front of the function name and click run, R will open a help page for the function.
# 
# This is the best way to learn about how the function works. Notice that you do not include parentheses when you open a help page.

```


### Run a function

Can you use the `sqrt()` function in the chunk below to compute the square root of 962?

::: {.panel-tabset}
## {{< fa code >}}&ensp;Interactive editor

```{webr}


```

## {{< fa circle-check >}}&ensp;Solution

```r
sqrt(962)
```

:::

### Code

Use the code chunk below to examine the code that `sqrt()` runs.

::: {.panel-tabset}
## {{< fa code >}}&ensp;Interactive editor

```{webr}



```

## {{< fa circle-check >}}&ensp;Solution

```r
sqrt
```

:::

###

Good job! `sqrt` immediately triggers a low level algorithm optimized for performance, so there is not much code to see.

### lm

Compare the code in `sqrt()` to the code in another R function, `lm()`. Examine `lm()`'s code body in the chunk below.

::: {.panel-tabset}
## {{< fa code >}}&ensp;Interactive editor

```{webr}



```

## {{< fa circle-check >}}&ensp;Solution

```r
lm
```

:::


### Help pages

Wow! `lm()` runs a lot of code. What does it do? Open the help page for `lm()` in the chunk below and find out.

::: {.panel-tabset}
## {{< fa code >}}&ensp;Interactive editor

```{webr}
#| autorun: false
?lm

```

## {{< fa circle-check >}}&ensp;Solution

```r
?lm
```

:::

###

Good job! `lm()` is R's function for fitting basic linear models. No wonder it runs so much code.


### Code comments

What do you think the chunk below will return? Run it and see. The result should be nothing. R will not run anything on a line after a `#` symbol. This is useful because it lets you write human readable comments in your code: just place the comments after a `#`. Now delete the `#` and re-run the chunk. You should see a result.

::: {.panel-tabset}
## {{< fa code >}}&ensp;Interactive editor

```{webr}
# sqrt(962)


```

## {{< fa circle-check >}}&ensp;Solution

```r
sqrt(962)
```

:::

##

```{r}
#| echo: false
#| results: asis
create_buttons("02-arguments.html")
```
