---
title: "Objects"
format:
  live-html: 
    toc: false
    section-divs: true
    include-after-body: 
      - text: |
          <script src="/js/progressive-reveal.js"></script>

engine: knitr
webr:
  cell-options:
    editor-font-scale: 0.85
    fig-width: 6
    fig-height: 3.7
    out-width: "70%"
---


{{< include /_extensions/r-wasm/live/_knitr.qmd >}}


```{r include=FALSE}
source(here::here("R", "helpers.R"))
```

### Objects {.no-hide}

So far we've been happy to run functions and then read the results on the screen. If you'd like to use those results later you'll need to save them.

In R you save results by creating an object. To create an object type a name for the object and then use the assignment operator to store a value in the object.

```{webr}
my_num <- 7
mynum
```


The operator is a less than sign followed by a minus sign. The result looks like an arrow which is a helpful visual mnemonic.

R will store whatever appears on the right of the arrow into the object that appears on the left. You can save individual values or you can directly save the results of a function.

```{webr}
my_log_result <- log(10,2)
my_log_result
```

Whenever you call an object R will replace the object with the value assigned to the object when R executes the code.

```{webr}
my_log_result
```

This makes it easy to pass results to new functions or to see the contents of an object. To do that just call the object name by itself.

```{webr}
my_num + my_log_result
```

If you assign a value to an object that already exists R won't hesitate to overwrite the old value. So be careful when you choose your object names.

```{webr}
my_num <- 999
my_num
```


### Object names

You can choose almost any name you like for an object, as long as the name does not begin with a number or a special character like `+`, `-`, `*`, `/`, `^`, `!`, `@`, or `&`. 

For instance, check out this list of some possible object names. Some are okay to use; some are invalid:

- `today`: This is fine
- `1st`: This is **bad**; it starts with a number
- `+1`: This is **bad**; it starts with a special character
- `vars`: This is fine
- `\^_^`: This is **bad**; it starts with a special character
- `foo`: This is fine


### Using objects

In the code chunk below, save the results of `rnorm(100, mean = 100, sd = 15)` to an object named `data`. Then, on a new line, call the `hist()` function on `data` to plot a histogram of the random values.

::: {.panel-tabset}
## {{< fa code >}}&ensp;Interactive editor

```{webr}



```

## {{< fa circle-check >}}&ensp;Solution

```r
data <- rnorm(100, mean = 100, sd = 15)
hist(data)
```

:::

### What if?

What do you think would happen if you assigned `data` to a new object named `copy`, like this? Run the code and then inspect both `data` and `copy`.

::: {.panel-tabset}
## {{< fa code >}}&ensp;Interactive editor

```{webr}
data <- rnorm(100, mean = 100, sd = 15)
copy <- data


```

## {{< fa circle-check >}}&ensp;Solution

```r
data <- rnorm(100, mean = 100, sd = 15)
copy <- data
data
copy
```

:::

###

Good job! R saves a copy of the contents in data to copy.

### Datasets

Objects provide an easy way to store datasets in R. In fact, R comes with many toy datasets pre-loaded. Examine the contents of `mtcars` to see a classic toy dataset. Hint: how could you learn more about the `mtcars` object?

::: {.panel-tabset}
## {{< fa code >}}&ensp;Interactive editor

```{webr}



```

## {{< fa circle-check >}}&ensp;Solution

```r
mtcars
```

:::

###

Good job! You can learn more about mtcars by examining its help page with `?mtcars`.

### `rm()`

What if you accidentally overwrite an object? If that object came with R or one of its packages, you can restore the original version of the object by removing your version with `rm()`. Run `rm()` on `mtcars` below to restore the mtcars data set.

::: {.panel-tabset}
## {{< fa code >}}&ensp;Interactive editor

```{webr}
mtcars <- 1
mtcars


```

## {{< fa circle-check >}}&ensp;Solution

```r
mtcars <- 1
mtcars
rm(mtcars)
mtcars
```

:::

###

Good job! Unfortunately, `rm()` cannot help you if you overwrite one of your own objects.

##

```{r}
#| echo: false
#| results: asis
create_buttons("04-vectors.html")
```
