---
title: "Vectors"
format:
  live-html: 
    toc: false
    section-divs: true
    include-after-body: 
      - text: |
          <script src="/js/progressive-reveal.js"></script>

engine: knitr
webr:
  cell-options:
    editor-font-scale: 0.85
    fig-width: 6
    fig-height: 3.7
    out-width: "70%"
---

{{< include /_extensions/r-wasm/live/_knitr.qmd >}}


```{r include=FALSE}
source(here::here("R", "helpers.R"))
```

### Vectors {.no-hide}

R stores every value as a vector, which is a one dimensional array of values. 
When you create a single value in R, R sees the value as a vector of length one.

When you run this `rnorm()` call, R sees the result as a vector of length ten. 

```{webr}
rnorm(10)
```

You can create your own vectors with the `c()` function which combines the values into a single vector.

```{webr}
c(1, 3, 5, 7, 9)
```

If you want to, you can assign each value a name as you do so. Creating your own vectors is a very useful thing to do because many R functions expect a vector.

```{webr}
c(a = 1, b = 3, c = 5, d = 7, e = 9)
```


For example, `mean()` computes the mean of a vector of values, and `sum()` computes the sum. Vectors also help you work efficiently. You can give many R functions a vector of input. R will apply the function separately to each element in the vector and return the result as a new vector.

If you use multiple vectors of input R will apply the function to the first 
element of each vector, and then to the second element of each vector, and 
so on until R reaches the end of the vector.

### Create a vector

In the chunk below, create a vector that contains the integers from one to ten. Use the `c()` function.

::: {.panel-tabset}
## {{< fa code >}}&ensp;Interactive editor

```{webr}



```

## {{< fa circle-check >}}&ensp;Solution

```r
c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10)
```

:::


### `:`

If your vector contains a sequence of contiguous integers, you can create it with the `:` shortcut. Run `1:10` in the chunk below. What do you get? What do you suppose `1:20` would return?

::: {.panel-tabset}
## {{< fa code >}}&ensp;Interactive editor

```{webr}



```

## {{< fa circle-check >}}&ensp;Solution

```r
1:10
1:20
```

:::


### `[]`

You can extract any element of a vector by placing a pair of brackets behind the vector. Inside the brackets place the number of the element that you'd like to extract. For example, `vec[3]` would return the third element of the vector named `vec`.

Use the chunk below to extract the fourth element of `vec`.

::: {.panel-tabset}
## {{< fa code >}}&ensp;Interactive editor

```{webr}
vec <- c(1, 2, 4, 8, 16)


```

## {{< fa circle-check >}}&ensp;Solution

```r
vec <- c(1, 2, 4, 8, 16)
vec[4]
```

:::

### More `[]`

You can also use `[]` to extract multiple elements of a vector. Place the vector `c(1,2,5)` between the brackets below. What does R return?

::: {.panel-tabset}
## {{< fa code >}}&ensp;Interactive editor

```{webr}
vec <- c(1, 2, 4, 8, 16)
vec[]


```

## {{< fa circle-check >}}&ensp;Solution

```r
vec <- c(1, 2, 4, 8, 16)
vec[c(1,2,5)]
```

:::


### Names

If the elements of your vector have names, you can extract them by name. To do so place a name or vector of names in the brackets behind a vector. Surround each name with quotation marks, e.g. `vec2[c("alpha", "beta")]`.

Extract the element named gamma from the vector below.

::: {.panel-tabset}
## {{< fa code >}}&ensp;Interactive editor

```{webr}
vec2 <- c(alpha = 1, beta = 2, gamma = 3)


```

## {{< fa circle-check >}}&ensp;Solution

```r
vec2 <- c(alpha = 1, beta = 2, gamma = 3)
vec2["gamma"]
```

:::


### Vectorised operations

Predict what the code below will return. Then look at the result.

::: {.panel-tabset}
## {{< fa code >}}&ensp;Interactive editor

```{webr}
c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10) + c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10)


```

:::

###

Good job! Like many R functions, R's math operators are vectorized: they're designed to work with vectors by repeating the operation for each pair of elements.

### Vector recycling

Predict what the code below will return. Then look at the result.

::: {.panel-tabset}
## {{< fa code >}}&ensp;Interactive editor

```{webr}
1 + c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10)


```

:::

###

Good job! Whenever you try to work with vectors of varying lengths (recall that `1` is a vector of length one), R will repeat the shorter vector as needed to compute the result.

##

```{r}
#| echo: false
#| results: asis
create_buttons("05-types.html")
```
