---
title: "Types"
format:
  live-html: 
    toc: false
    section-divs: true
    include-after-body: 
      - text: |
          <script src="/js/progressive-reveal.js"></script>
          <script src="/js/bootstrapify.js"></script>

engine: knitr
webr:
  cell-options:
    editor-font-scale: 0.85
    fig-width: 6
    fig-height: 3.7
    out-width: "70%"
---

{{< include /_extensions/r-wasm/live/_knitr.qmd >}}

```{r include=FALSE}
library(tidyverse)
library(checkdown)

source(here::here("R", "helpers.R"))
```

### Types {.no-hide}

R recognizes different types of vectors based on clues in your values. For 
example, if your values are all numbers, R knows that the vector is numeric 
and R can do math with it.

```{r}
mynums <- c(1,2,3,4)
sum(mynums)
```

If your values are surrounded by quotation marks, R knows that the vector contains character strings and R can do things like count the number of characters in each string.

```{r}
mywords <- c("cat", "zebra", "hippopotamus")
nchar(mywords)
```

By recognizing what type of data you're using, R can behave more intuitively. Altogether R recognizes six different types of data.

You can check the type of a vector with `typeof()`.

```{r}
typeof(mywords)
typeof(mynums)
```

Numeric vectors contain numbers and possibly decimal points and negative signs. They have the type double which is a computer science term that R associates with decimal numbers. Character vectors contain values surrounded by quotation marks.

Integer vectors contain integers followed by a capital L. Logical vectors contain 
only TRUEs and FALSEs written in all capital letters.

R also recognizes complex vectors and raw vectors, but we won't cover these because you are unlikely to encounter them while doing data science.

It is helpful to know that each vector in R can only contain one type of data. If you try to combine types within a single vector R will collapse everything to the same type, often a character vector.

If you want to save different types of values in the same object you will need to use a list, which is a data structure that we'll look at in the next video.

Later you'll learn about other classes of data that are built on top of R's basic types. R developers have come up with ingenious ways to use the six basic types to represent other classes of data like dates and times, and factors.

But for now, it's important to realize that every value in R is based on a double, character, integer, logical, complex, or raw vector.

### Atomic types

::: {.callout-note appearance="simple" icon=false .question}

**Which of these is not an atomic data type?**

```{r types1, echo=FALSE}
check_question(
  answer = "simple", 
  options = c(
   "numeric/double",
   "integer",
   "character",
   "logical",
   "complex",
   "raw",
   "simple"
  ), 
  type = "radio", 
  button_label = "Submit answer", 
  q_id = 1,
  right = c("Correct!")
)
```

:::

### What type?

::: {.callout-note appearance="simple" icon=false .question}

**What type of data is `"1L"`?**

```{r types2, echo=FALSE}
check_question(
  answer = "character", 
  options = c(
   "numeric/double",
   "integer",
   "character",
   "logical"
  ), 
  type = "radio", 
  button_label = "Submit answer", 
  q_id = 2,
  right = c("Correct! This was tricky because of the quotes. 1L by itself would be an integer, but values become characters when they're in quotes.")
)
```

:::

### Integers

Create a vector of integers from one to five. Can you imagine why you might want to use integers instead of numbers/doubles?

::: {.panel-tabset}
## {{< fa code >}}&ensp;Interactive editor

```{webr}



```

## {{< fa circle-check >}}&ensp;Solution

```r
c(1L, 2L, 3L, 4L, 5L)
```

:::


### Floating point arithmetic

Computers must use a finite amount of memory to store decimal numbers (which can sometimes require infinite precision). As a result, some decimals can only be saved as very precise approximations. From time to time you'll notice side effects of this imprecision, like below.

Compute the square root of two, square the answer (e.g. multiply the square root of two by the square root of two), and then subtract two from the result. What answer do you expect? What answer do you get?

::: {.panel-tabset}
## {{< fa code >}}&ensp;Interactive editor

```{webr}



```

## {{< fa circle-check >}}&ensp;Solution

```r
sqrt(2) * sqrt(2) - 2
sqrt(2)^2 - 2
```

:::


### Vectors

::: {.callout-note appearance="simple" icon=false .question}

**How many types of data can you put into a single vector?**

```{r types3, echo=FALSE}
check_question(
  answer = "1", 
  options = c(
   "1",
   "6",
   "As many as you like"
  ), 
  type = "radio", 
  button_label = "Submit answer", 
  q_id = 3,
  right = c("Correct!")
)
```

:::

### Character or object?

One of the most common mistakes in R is to call an object when you mean to call a character string and vice versa.

::: {.callout-note appearance="simple" icon=false .question}

**Which of these are object names? What is the difference between object names and character strings?**

```{r types4, echo=FALSE}
check_question(
  answer = c("foo", "mu", "a"), 
  options = c(
   "foo",
   '"num"',
   "mu",
   '"sigma"',
   '"data"',
   "a"
  ), 
  type = "checkbox", 
  button_label = "Submit answer", 
  q_id = 4,
  right = c("Correct! Character strings are surrounded by quotation marks, object names are not.")
)
```

:::


##

```{r}
#| echo: false
#| results: asis
create_buttons("06-lists.html")
```
