---
title: "Lists"
format:
  live-html: 
    toc: false
    section-divs: true
    include-after-body: 
      - text: |
          <script src="/js/progressive-reveal.js"></script>
          <script src="/js/bootstrapify.js"></script>

engine: knitr
webr:
  cell-options:
    editor-font-scale: 0.85
    fig-width: 6
    fig-height: 3.7
    out-width: "70%"
---

{{< include /_extensions/r-wasm/live/_knitr.qmd >}}

```{r include=FALSE}
library(tidyverse)
library(checkdown)

source(here::here("R", "helpers.R"))
```

### Lists {.no-hide}

In addition to vectors, R comes with a second very useful data structure: 
the list. A list is like a vector of vectors. Each element in the list can be a separate vector or even another list.

For example, this list contains three elements.

```{r}
list(c(1,2,3,4,5), c("cat"), c(TRUE))
```

The first element is a numerical vector length five. The second is a character vector of length one, and the third element is a logical vector of length one. This list illustrates a useful feature of lists. You can store different types of data in the same list because each element of the list is its own vector with its own type.

Recall that basic vectors can only store a single type of data. You can make lists like this with the `list()` function.

`list()` works like the `c()` function. Give list the elements of the list to make, separated by commas. If you like you can assign names to the elements as you go (which I recommend).

If you assign names to each element of the list you can access the contents of the elements by typing the name of the list followed by a dollar sign, followed by the name of the element.

Lists are essential to data science because R uses a specific type of list to store tabular data, and that's the data frame.

R displays data frames (like the `penguins` data frame here) as a table. But R saves data frames as a list. Each column of the data frame is a vector that is saved as an element of the list.

```{r}
penguins
```

```{r}
#| include: false
# For example we can see how R saves `penguins`. It is actually a list of five vectors. The names of the columns in iris are the names of the elements in the list.

```

### Lists vs. vectors

::: {.callout-note appearance="simple" icon=false .question}

**Which data structure(s) could you use to store these pieces of data in the same object? `1001`, `TRUE`, `"stories"`**

```{r lists1, echo=FALSE}
check_question(
  answer = c("a list"), 
  options = c(
   "a vector",
   "a list",
   "neither"
  ), 
  type = "radio", 
  button_label = "Submit answer", 
  q_id = 1,
  right = c("Correct! Lists can contain elements that are different types.")
)
```

:::


### Make a list

Make a list that contains the elements `1001`, `TRUE`, and `"stories"`. Give each element a name.

::: {.panel-tabset}
## {{< fa code >}}&ensp;Interactive editor

```{webr}



```

## {{< fa circle-check >}}&ensp;Solution

```r
list(number = 1001, logical = TRUE, string = "stories")
```

:::


### Extract an element

Extract the number 1001 from the list below.

::: {.panel-tabset}
## {{< fa code >}}&ensp;Interactive editor

```{webr}
things <- list(number = 1001, logical = TRUE, string = "stories")


```

## {{< fa circle-check >}}&ensp;Solution

```r
things <- list(number = 1001, logical = TRUE, string = "stories")
things$number
```

:::

### Data Frames

You can make a data frame with the `data.frame()` function, which works similar to `c()`, and `list()`. Assemble the vectors below into a data frame with the column names `numbers`, `logicals`, `strings`.

::: {.panel-tabset}
## {{< fa code >}}&ensp;Interactive editor

```{webr}
nums <- c(1, 2, 3, 4)
logs <- c(TRUE, TRUE, FALSE, TRUE)
strs <- c("apple", "banana", "carrot", "duck")


```

## {{< fa circle-check >}}&ensp;Solution

```r
nums <- c(1, 2, 3, 4)
logs <- c(TRUE, TRUE, FALSE, TRUE)
strs <- c("apple", "banana", "carrot", "duck")
data.frame(numbers = nums, logicals = logs, strings = strs)
```

:::

###

Good job. When you make a data frame, you must follow one rule: each column vector should be the same length


### Extract a column

Given that a data frame is a type of list (with named elements), how could you extract the strings column of the `df` data frame below? Do it.

::: {.panel-tabset}
## {{< fa code >}}&ensp;Interactive editor

```{webr}
nums <- c(1, 2, 3, 4)
logs <- c(TRUE, TRUE, FALSE, TRUE)
strs <- c("apple", "banana", "carrot", "duck")
df <- data.frame(numbers = nums, logicals = logs, strings = strs)


```

## {{< fa circle-check >}}&ensp;Solution

```r
df$strings
```

:::

##

```{r}
#| echo: false
#| results: asis
create_buttons("07-packages.html")
```
